An Aliasing-Tolerant Parallel Programming Model

### 2012-11-09 doc/types-pldi2013/alt_example.tex

METIS takes as arguments lists of the nodes and wires in
the graph.  It breaks the graph into pieces by annotating each 
{\tt CircuitNode} with a piece ID (the last field of the tuple).
The function {\tt extern\_metis} is parameterized on the regions it 
accesses: {\tt rn} is a region of nodes, {\tt rw} is a region of wires, 
and {\tt rl} is a region with the lists of pointers.  The function
also specifies the access {\em privileges} (e.g read, write) it requires for each 
region.  All three regions are read by {\tt extern\_metis}, but only the {\tt rn}
region is written (since it writes the piece ID).  A function can only 
have privileges that its caller also possesses; this property is verified statically. 
% by the Legion type system.

@@

Programmers exoress partitions using {\em colorings},
maps from region elements to integers (``colors'').
Listing~\ref{lst:coloring} shows the function {\tt create\_piece\_coloring}
that constructs a coloring from the piece IDs of a list of nodes.
%The coloring of wires is similar, with each wire
%using the color of its first endpoint.  
For aliased partitions such as the ghost nodes, a {\em multicoloring}
is used, which assigns each region element (possibly) multiple colors.  
%(Multicolorings are
%implemented in Core Legion with separate colorings for each color.)
\begin{lstlisting}[label={lst:coloring},caption={Coloring Construction}]
function create_piece_coloring[rl,rn](node_list: NodeList<rl,rn>@rl),
            reads(rl,rn) : coloring(rn) = 
    if isnull(node_list) then
        newcolor(rn) 
    else
        let list_elem : NodeList<rl,rn> = read(node_list) in
        let part_coloring : coloring(rn) = create_node_coloring(list_elem.2) in
        let node : CircuitNode = read(list_elem.1) in
            color(part_coloring, list_elem.1, node.5)
\end{lstlisting}

Colorings are the input to partition operations in Legion.  The partition
statement shown in Listing~\ref{lst:partition} performs the ``owned nodes''
partitioning.    Since colorings assign each region element at most one color
(dynamically checked), the partition operation results in 
{\em disjoint} subregions, each a subset of the {\em parent} region.
A partition operation introduces constraints into the static type environment describing
both the disjointness of subregions (e.g., {\tt rn0 $*$ rn1})
and the subregion relationships
(e.g., {\tt rn0 $\leq$ all\_nodes}).  A partition created from a multicoloring
only introduces the subregion relationships, as the parition may be aliased.

@@

The Legion type system verifies properties of region relationships when packing.  When
a region relationship is unpacked later, fresh names are given to the regions that were
captured by the region relationship, and the properties that are known to hold for these
regions are re-introduced to the type environment.
In constrast, region access privileges cannot be packed in a region
relationship---privileges belong to functions. When a function unpacks
a region relationship it must already hold privileges for the unpacked regions it wishes to access.

>>>>

METIS takes as arguments lists of the nodes and wires in
the graph.  It breaks the graph into pieces by annotating each 
{\tt CircuitNode} with a piece ID (the last field of the tuple).
The function {\tt extern\_metis} is parametrized on the regions it 
accesses: {\tt rn} is a region of nodes, {\tt rw} is a region of wires, 
and {\tt rl} is a region with the lists of pointers.  The function
also specifies the access {\em privileges} (e.g read, write) it requires for each 
region.  All three regions are read by {\tt extern\_metis}, but only the {\tt rn}
region is written (since it writes the piece ID).  A function can only 
have privileges that its caller also possesses; this property is verified statically. 
% by the Legion type system.

@@

Programmers express partitions using {\em colorings},
maps from region elements to integers (``colors'').
Listing~\ref{lst:coloring} shows the function {\tt create\_piece\_coloring}
that constructs a coloring from the piece IDs of a list of nodes.
%The coloring of wires is similar, with each wire
%using the color of its first endpoint.  
For aliased partitions such as the ghost nodes, a {\em multicoloring}
is used, which assigns each region element (possibly) multiple colors.  
%(Multicolorings are
%implemented in Core Legion with separate colorings for each color.)
\begin{lstlisting}[label={lst:coloring},caption={Coloring Construction}]
function create_piece_coloring[rl,rn](node_list: NodeList<rl,rn>@rl),
            reads(rl,rn) : coloring(rn) = 
    if isnull(node_list) then
        newcolor(rn) 
    else
        let list_elem : NodeList<rl,rn> = read(node_list) in
        let part_coloring : coloring(rn) = create_node_coloring(list_elem.2) in
        let node : CircuitNode = read(list_elem.1) in
            color(part_coloring, list_elem.1, node.5)
\end{lstlisting}

Colorings are the input to partition operations in Legion.  The partition
statement shown in Listing~\ref{lst:partition} performs the ``owned nodes''
partitioning.    Since colorings assign each region element at most one color
(dynamically checked), the partition operation results in 
{\em disjoint} subregions, each a subset of the {\em parent} region.
A partition operation introduces constraints into the static type environment describing
both the disjointness of subregions (e.g., {\tt rn0 $*$ rn1})
and the subregion relationships
(e.g., {\tt rn0 $\leq$ all\_nodes}).  A partition created from a multicoloring
only introduces the subregion relationships, as the partition may be aliased.

@@

The Legion type system verifies properties of region relationships when packing.  When
a region relationship is unpacked later, fresh names are given to the regions that were
captured by the region relationship, and the properties that are known to hold for these
regions are re-introduced to the type environment.
In contrast, region access privileges cannot be packed in a region
relationship---privileges belong to functions. When a function unpacks
a region relationship it must already hold privileges for the unpacked regions it wishes to access.
